\papertitle

### 2012-10-11 manuscript.tex
final commit of the night

One characteristic of the H$^2$Bird that separates it from other MAVs is the
high pitch angles experienced in routine flight. An Euler angle 
parameterization of the vehicle orientation would suffer from singularities
around these high pitch angles, so we instead represent the vehicle orientation
with quaternions. An additional benefit of the quaternion representation is 
the ability to represent relative body and world coordinate angular displacements 
compactly with quaternion multiplication.~\cite{bowman:reasoning}

@@

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. The results of this experiment are depicted in
Figure~\ref{fig:step_response}. The data for the experiment was gathered from
the integration of the H$^2$Bird's built in gyroscope data. We used Matlab to
fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
turning behavior to the step response with the physical heading as the state
and the heading set point as the input. This model serves as a simplification
of both the turning dynamics of the ornithopter and the on-board PID
controller regulating the actuator inputs to reach the desired orientation. We
use this model in Section~\ref{sec:visual_servoing} to determine the backwards
reachable region for navigating through the window in simulation.
\\
The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
flies at an average forward velocity of 1.2 m/s, so the estimated minimum
turning radius of the robot is 1.07 meters. This flight speed is slower than
the maximum speed of the ornithopter to facilitate more reliable tracking.
%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
\label{sec:visual_servoing}
%TODO cameron

%TODO make these span the page, with identical scales/style
\begin{figure*}[h]

\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
\caption{Plot of backwards reachable set for successful window traversal.}
\label{fig:feasible_set}
\end{minipage}

\hspace{0.5cm}
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/flight_paths.pdf}
\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius 
cone, with experimental trials overlayed.}
\label{fig:flight_paths}
\end{minipage}

\end{figure*}

@@

We determined the backwards reachable set of initial states for successful
window traversal both geometrically and in simulation. The results are shown
in Figure~\ref{fig:feasible_set}. The initial conditions that we chose are
spatially gridded in the camera viewing region at 0.05 meter increments and an
initial angular position perpendicular to the window plane. The geometrically
infeasible region is indicated in blue and was computed using the minimum
turning radius and simple geometry. Both the regions in red and blue indicate
the infeasible region determined in simulation. We simulated the turning model
described in Section~\ref{sec:flight_control}, with a PID controller steering
the system to the center of the window as the input. The infeasible region
calculated in simulation is a superset of the geometrically infeasible region.
The simulation is more realistic as the rotational damping during the turn and
the on-board PID controller cause the real system to react to inputs at a
non-constant angular velocity.
\\
To verify the feasible region experimentally, we conducted trials at initial
distances in approximately 0.4 meter increments from the window along the
edges of the image frame of the camera. We started each trial with an initial
heading perpendicular to the window plane. The results of the experiments are
shown in Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.

For initial conditions in the middle of the feasible region and in front of
the camera, we achieved a success rate of 79\%. This case is depicted in
Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
feasible region (Figure~\ref{flight_paths}), the success rate diminishes
because of several factors. On the edges of the camera frame, the yaw control
inputs are large in magnitude, although this is not necessary in all cases.
Since the control algorithm we are using has no notion of depth, it applies
the same control input whether the ornithopter is close to the camera, where
small changes in heading are adequate, or far away from the camera, where
large changes in heading are necessary. Due to this phenomenon, the controller
often over- or under-compensates in certain regions, depending on the PID
tuning and the distance from the camera.

>>>>

One characteristic of the H$^2$Bird that separates it from other MAVs is the
high pitch angles experienced in routine flight. An Euler angle 
parameterization of the vehicle orientation would suffer from singularities
around these high pitch angles, so we instead represent the vehicle orientation
with quaternions. An additional benefit of the quaternion representation is 
the ability to represent relative body and world coordinate angular displacements 
compactly with quaternion multiplication.%~\cite{bowman:reasoning}

@@

We determined the turning speed and turning radius of the H$^2$Bird
experimentally using the response of the system to the step input of a
clockwise 90 degree turn. The results of this experiment are depicted in
Figure~\ref{fig:step_response}. The data for the experiment was gathered from
the integration of the H$^2$Bird's on-board gyroscope data. We used Matlab to
fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
turning behavior to the step response with the physical heading as the state
and the heading set point as the input. This model serves as a simplification
of both the turning dynamics of the ornithopter and the on-board PID
controller regulating the actuator inputs to reach the desired orientation. We
use this model in Section~\ref{sec:visual_servoing} to determine the backwards
reachable region for navigating through the window in simulation.
\\
The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
flies at an average forward velocity of 1.2 m/s, so the estimated minimum
turning radius of the robot is 1.07 meters. This flight speed is slower than
the maximum speed of the ornithopter to facilitate more reliable tracking.
%----------------------------------------------------------------------------%
\subsection{Visual Servoing}
\label{sec:visual_servoing}
%TODO cameron

%TODO make these span the page, with identical scales/style
\begin{figure*}[tb]
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
\caption{Plot of backwards reachable set for successful window traversal.}
\label{fig:feasible_set}
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/flight_paths.pdf}
\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius 
cone, with experimental trials overlayed.}
\label{fig:flight_paths}
\end{minipage}
\end{figure*}

@@

We determined the backwards reachable set of initial states for successful
window traversal both geometrically and in simulation. The results are shown
in Figure~\ref{fig:feasible_set}. For the initial conditions, we used a
spatial grid of 0.05 meter increments within the camera viewing region and an
initial angular position perpendicular to the window plane. The geometrically
infeasible region is indicated in blue and was computed using the minimum
turning radius and simple geometry. Both the regions in red and blue indicate
the infeasible region determined in simulation. We simulated the turning
model, described in Section~\ref{sec:flight_control}, with a PID controller
steering the system to the center of the window as the input. The infeasible
region calculated in simulation is a superset of the geometrically infeasible
region. The simulation is more realistic as the rotational damping during the
turn and the on-board PID controller cause the real system to react to inputs
at a non-constant angular velocity.
\\
To verify the feasible region experimentally, we conducted trials at initial
distances in approximately 0.4 meter increments from the window along the
edges of the camera image frame. We began each trial with an initial heading
perpendicular to the window plane. The results of the experiments are shown in
Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
\\
For initial conditions in the middle of the feasible region and in front of
the camera, we achieved a success rate of 79\%. This case is depicted in
Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
because of several factors. On the edges of the camera frame, the yaw control
inputs are large in magnitude, although this is not necessary in all cases.
Since the control algorithm we are using has no notion of depth, it applies
the same control input whether the ornithopter is close to the camera, where
small changes in heading are adequate, or far away from the camera, where
large changes in heading are necessary. Due to this phenomenon, the controller
often over- or under-compensates in certain regions, depending upon the PID
tuning and the distance from the camera. Tracking noise can also cause the
window navigation to fail.
